# Different ways to Impute Missing Values

[Original Article](https://appsilon.com/imputation-in-r/#:~:text=Summary-,Introduction%20to%20Imputation%20in%20R,into%20a%20machine%20learning%20model)


Required packages:
```{r}
library(ggplot2)
library(dplyr)
library(cowplot)
library(titanic)
```

Looking at the age data overall first to get a sense of what it already is
```{r}
head(titanic_train$Age)

ggplot(titanic_train, aes(Age))+
    geom_histogram(color="black",fill="skyblue")+
    ggtitle("Variable distribution")+
    theme_classic()+
    theme(plot.title = element_text(size = 18))

```


### Simple Value Imputation using built in functions

- Zero: a constant put in for all NAs
- Mean (average): average value for age replacing all NAs
- Median: median value for age replacing all NAs

```{r}
value_imputed<- data.frame(
    original=titanic_train$Age,
    imputed_zero=replace(titanic_train$Age, is.na(titanic_train$Age),0),
    imputed_mean=replace(titanic_train$Age, is.na(titanic_train$Age),mean(titanic_train$Age, na.rm = TRUE)),
    imputed_median=replace(titanic_train$Age, is.na(titanic_train$Age),median(titanic_train$Age, na.rm = TRUE))
)
head(value_imputed, n=15)
```

Look at how each type of imputed data changed the distribution:
```{r}
h1 <- ggplot(value_imputed, aes(x = original)) +
  geom_histogram(fill = "#ad1538", color = "#000000", position = "identity") +
  ggtitle("Original distribution") +
  theme_classic()
h2 <- ggplot(value_imputed, aes(x = imputed_zero)) +
  geom_histogram(fill = "#15ad4f", color = "#000000", position = "identity") +
  ggtitle("Zero-imputed distribution") +
  theme_classic()
h3 <- ggplot(value_imputed, aes(x = imputed_mean)) +
  geom_histogram(fill = "#1543ad", color = "#000000", position = "identity") +
  ggtitle("Mean-imputed distribution") +
  theme_classic()
h4 <- ggplot(value_imputed, aes(x = imputed_median)) +
  geom_histogram(fill = "#ad8415", color = "#000000", position = "identity") +
  ggtitle("Median-imputed distribution") +
  theme_classic()

plot_grid(h1, h2, h3, h4, nrow = 2, ncol = 2)
```

### Using MICE

Load in MICE and look at the numeric columns
```{r}
library(mice)

titanic_numeric<-titanic_train %>%
    select(Survived, Pclass, SibSp, Parch, Age)

md.pattern(titanic_numeric)
```

Use these three types of imputation methods from MICE:

- pmm: predictive mean matching
- cart: classification and regression trees
- laso.norm: lasso linear regression

```{r}

mice_imputed <- data.frame(
  original = titanic_train$Age,
  imputed_pmm = complete(mice(titanic_numeric, method = "pmm"))$Age,
  imputed_cart = complete(mice(titanic_numeric, method = "cart"))$Age,
  imputed_lasso = complete(mice(titanic_numeric, method = "lasso.norm"))$Age
)
head(mice_imputed, n=15)
```

Look at different distributions again:

```{r}
s1 <- ggplot(mice_imputed, aes(x = original)) +
  geom_histogram(fill = "#ad1538", color = "#000000", position = "identity") +
  ggtitle("Original distribution") +
  theme_classic()
s2 <- ggplot(mice_imputed, aes(x = imputed_pmm)) +
  geom_histogram(fill = "#15ad4f", color = "#000000", position = "identity") +
  ggtitle("PMM-imputed distribution") +
  theme_classic()
s3 <- ggplot(mice_imputed, aes(x = imputed_cart)) +
  geom_histogram(fill = "#1543ad", color = "#000000", position = "identity") +
  ggtitle("Cart-imputed distribution") +
  theme_classic()
s4 <- ggplot(mice_imputed, aes(x = imputed_lasso)) +
  geom_histogram(fill = "#ad8415", color = "#000000", position = "identity") +
  ggtitle("Lasso-imputed distribution") +
  theme_classic()

plot_grid(s1, s2, s3, s4, nrow = 2, ncol = 2)
```

### Using the missForest package

```{r}
library(missForest)

missForest_imputed <- data.frame(
  original = titanic_numeric$Age,
  imputed_missForest = missForest(titanic_numeric)$ximp$Age
)
head(missForest_imputed, n=15)

```

```{r}

f1 <- ggplot(missForest_imputed, aes(x = original)) +
  geom_histogram(fill = "#ad1538", color = "#000000", position = "identity") +
  ggtitle("Original distribution") +
  theme_classic()
f2 <- ggplot(missForest_imputed, aes(x = imputed_missForest)) +
  geom_histogram(fill = "#15ad4f", color = "#000000", position = "identity") +
  ggtitle("missForest-imputed distribution") +
  theme_classic()

plot_grid(f1, f2, nrow = 1, ncol = 2)
```













